---
title: "MATH 216 Homework 3"
author: "Connor"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    smooth_scroll: false
---

```{r, echo=FALSE, message=FALSE}
library(ggplot2)
library(dplyr)
library(readr)
library(lubridate)
library(Quandl)
library(forcats)
library(tidyverse)
library(norm)
library(knitr)
```



## Admistrative:

Please indicate

* Who you collaborated with: Kyra, Trisha
* Roughly how much time you spent on this HW so far: 9
* The URL of the RPubs published URL [here](http://rpubs.com/connormccormick/hw3).
* What gave you the most trouble: Dealing with the separted variables in the jukebox dataframe
* Any comments you have:




## Question 1:

We will use a logistic regression model to predict sex. Our metric to rate how well our
model performs will be:

$$
\frac{1}{n}\sum_{i=1}^{n}I(y_i = \widehat{y}_i)
$$

where $I(A)$ is the *indicator function* that is equal to 1 if condition $A$
holds, 0 otherwise. So

* Say user $i$ **is** female, then $y_i=1$
* Say we **predict** user $i$ is female, then $\widehat{y}_i=1$
* In this case $I(y_i =\widehat{y}_i)=1$. 

So what the above formula is reporting is the proportion of users' sex we
correctly predicted.

```{r, echo=FALSE, message=FALSE, cache=TRUE}
# Edit this code block at your own peril! cache is set to TRUE!
# To keep this exercise simple, let's remove the only 3 users (0.005% of users)
# who did not list a height, define the outcome variable, and add an ID variable
# to distinguish the users
profiles <- read_csv(file="profiles.csv") %>% 
  filter(!is.na(height)) %>% 
  mutate(is_female=ifelse(sex=='f', 1, 0)) %>% 
  tibble::rownames_to_column(var="id")
```



#### a)

Define:

* A *training* set `training` of 2997 users (5% of users). We will train the 
logistic regression model to predict gender using this data. Since we want to 
train the model to tell who is female and who is not, we use the outcome
variable `is_female`.
* A *test* set `test` of the remaining 56,946 users (95% of users). We will test
how good our trained model is using this data. So at first, we will pretend we
don't know the outcome variable `is_female`. We use the above model to make a
prediction of sex for all 56,946 test users, then we use the `is_female` outcome
to rate how well we performed.
* Be sure to incorporate all the insight your garnered in your EDA in HW-2.

```{r, echo=TRUE, message=TRUE, warning=TRUE, fig.width=8, fig.height=4.5}
cleanprofiles <- profiles %>%
  select(id, sex, income, job, pets, is_female) %>%
  mutate(petowners = fct_recode(pets,
                            # New name = old name
                           "dogs" = "has dogs",
                           "dogs" = "has dogs and dislikes cats",
                           "dogs" = "has dogs and likes cats",
                           "both" = "has dogs and has cats",
                           "cats" = "has cats",
                           "cats" = "dislikes dogs and has cats",
                           "cats" = "likes dogs and has cats",
                           "neither" = "dislikes cats",
                           "neither" = "dislikes dogs",
                           "neither" = "dislikes dogs and dislikes cats",
                           "neither" = "dislikes dogs and likes cats",
                           "neither" = "likes cats",
                           "neither" = "likes dogs",
                           "neither" = "likes dogs and dislikes cats",
                           "neither" = "likes dogs and likes cats"
  )) 
  

cleaner <- cleanprofiles %>%
  mutate(job = fct_recode(job, 
                          "other" = "military",
                          "other" = "unemployed",
                          "other" = "transportation",
                          "other" = "retired",
                          "other" = "rather not say",
                          "other" = "political / government",
                          "other" = "clerical / administrative",
                          "other" = "hospitality / travel",
                          "other" = "law / legal services",
                          "other" = "construction / craftsmanship"
    ))

newclean <- cleaner %>%
  select(id, income, job, is_female, petowners) %>%
  mutate(petowners = ifelse(is.na(petowners), "N/A", as.character(petowners))) %>%
  mutate(job = ifelse(is.na(job), "N/A", as.character(job)))

set.seed(21)
training <- sample_n(newclean, 2997)
test <- newclean %>%
  filter(!(id %in% training$id))

#graph to show proportion by sex
gender <- training %>%
  group_by(is_female) %>%
  tally()%>%
  rename(gender = n) %>%
  mutate(prop = gender/sum(gender))

ggplot(gender, aes(x = as.factor(is_female), y = prop)) +
  geom_bar(stat = "identity", position = "dodge", fill = "violetred4", color = "darkslateblue", width = .65) +
  theme_minimal() +
  labs(title = "Proportion by Sex of Sample", x = "Female", y = "Proportion")  

```



#### b)

Train the logistic regression model to predict sex. i.e. fit a logistic
regression model to the `training` data. Assign this model to an R object called
`predict_sex_model`, then rate how well the model performs on the `training` data.

```{r, echo=TRUE, message=TRUE, warning=TRUE, fig.width=8, fig.height=4.5}
#regression
predictsexmodel <- glm(is_female ~ income + job + petowners, data=training, family="binomial")

predictions <- training %>%
  select(income, job, petowners, is_female) %>%
  mutate(p_hat1 = fitted(predictsexmodel))


#How did the model do?
prophecy <- predictions %>%
  mutate(guess = ifelse(p_hat1>=.5, 1, 0))

errors <- prophecy %>%
  mutate(correct = ifelse(guess==is_female, 1, 0))


propwrong <- errors %>%
  summarize(prop_wrong=1 - mean(correct))
kable(propwrong)

sexpropwrong <- errors %>%
  group_by(is_female) %>%
  summarize(prop_wrong=1-mean(correct))
kable(sexpropwrong)

```

The first table looks at the proportion of wrong answers generated by the model for the training data. 35.40% of the time, the model guest the wrong sex for a person based on their OkCupid profile.
The second table broke down the incorrect answers by sex, with female=1.  The model was only wrong 18.90% of the time when the respondent was male, compared to being wrong 58.59% of the time when the respondent was female.  


#### c)

Take `predict_sex_model` and apply it to the `test` data and make a prediction 
for each users' sex, then rate how well the model performs on the `test` data.

**Hint**: What do you think `predict(predict_sex_model, newdata=test,
type="response")` does? The help file is located in `?predict.glm`

```{r, echo=TRUE, message=TRUE, warning=TRUE, fig.width=8, fig.height=4.5}
test <- test %>% 
  select(is_female, income, job, petowners) %>%
  mutate(phat=predict(predictsexmodel, newdata = test, type = "response"))

#How did the model do for the test?
testprophecy <- test %>%
  mutate(guess = ifelse(phat>=.5, 1, 0))

testerrors <- testprophecy %>%
  mutate(correct = ifelse(guess==is_female, 1, 0))
  
testpropwrong <- testerrors %>%
  summarize(prop_wrong=1 - mean(correct))
kable(testpropwrong)

testsexpropwrong <- testerrors %>%
  group_by(is_female) %>%
  summarize(prop_wrong=1-mean(correct))
kable(testsexpropwrong)


```


The first table looks at the proportion of wrong answers generated by the model for the test data. 35.39% of the time, the model guest the wrong sex for a person based on their OkCupid profile.
The second table broke down the incorrect answers by sex, with female=1.  The model was only wrong 18.54% of the time when the respondent was male, compared to being wrong 60.49% of the time when the respondent was female.  


#### d)

Did the model perform better on the `training` data or the `test` data? Why
do you think that is?

The model performed marginally better in the test data than in the training data.  The model was wrong 35.40% of the time in the training compared to 35.39% in the test. Surprisingly, the model performed better for men in the test and better for women in the training.   





## Question 2:

We want to compare the volatility of 

* [Bitcoin](https://www.quandl.com/data/BAVERAGE/USD) prices
* [Gold](https://www.quandl.com/data/BUNDESBANK/BBK01_WT5511) prices

Let our measure of volatility be the relative change from day-to-day in price. 
Let the reference currency be US dollars. Analyze these results and provide
insight to a foreign currency exchanger.

```{r, echo=TRUE, message=TRUE, warning=TRUE, fig.width=8, fig.height=4.5}
bitcoin <- Quandl("BAVERAGE/USD") %>%
  tbl_df() %>%
  rename(
    Avg = `24h Average`,
    Total_Volume = `Total Volume`
  ) 

gold <- Quandl("BUNDESBANK/BBK01_WT5511")

bitcoin$lagged <- lead(bitcoin$Avg, 1)

golddates <- interval(ymd("2010-07-17"), ymd("2016-04-15"))

bitcoin <-bitcoin %>%
  filter(!is.na(lagged)) %>%
  filter(Date %within% golddates) %>%
  mutate(lagvalue = (Avg - lagged)/lagged)


gold$lagged <- lead(gold$Value, 1)

bitcoindates <- interval(ymd("2010-07-17"), ymd("2016-10-22"))

gold <-gold %>%
  filter(!is.na(lagged)) %>%
  filter(Date %within% bitcoindates) %>%
  mutate(lagvalue = (Value - lagged)/lagged) 

  

ggplot(bitcoin, aes(x = Date, y = lagvalue)) +
  geom_line(color = "darkslateblue") +
  geom_line(data = gold, color = "cadetblue2", alpha = .75) +
  theme_minimal() +
  labs(title = "Daily Volatility of Bitcoin and Gold", y = "Percent Daily Change")




```
Bitcoin is substantially more volatile than gold when looking at percent daily changes.  




## Question 3:

```{r, echo=FALSE, message=FALSE, cache=TRUE}
# Edit this code block at your own peril! cache is set to TRUE!
jukebox <- read_csv(file="reed_jukebox.csv")

# Clean certain artists' names:
sigur_ros <- "Sigur Ro\xfc\xbe\x99\x86\x94\xbc\xfc\xbe\x8c\x93\xa0\xbcs"
bjork <- "Bjo\xfc\xbe\x99\x86\x94\xbc\xfc\xbe\x8d\xa6\x98\xbcrk"
blue_oyster_cult <- "Blue O\xfc\xbe\x99\x86\x94\xbc\xfc\xbe\x8d\xa6\x98\xbcyster Cult"
husker_do <- "Hu\xfc\xbe\x99\x86\x94\xbc\xfc\xbe\x8d\xa6\x98\xbcsker Du\xfc\xbe\x99\x86\x94\xbc\xfc\xbe\x8d\xa6\x98\xbc"
bjork_brodsky <- "Bjo\xfc\xbe\x99\x86\x94\xbc\xfc\xbe\x8d\xa6\x98\xbcrk & Brodsky Quartet"
slagsmalsklubben <- "Slagsma\xfc\xbe\x99\x86\x94\xbc_lsklubben "
bjork_sugarcubes <- "Bjo\xfc\xbe\x99\x86\x94\xbc\xfc\xbe\x8d\xa6\x98\xbcrk (Sugarcubes)"

jukebox <- jukebox %>%
  mutate(
    artist = ifelse(artist == sigur_ros, "Sigor Ros", artist),
    artist = ifelse(artist == bjork, "Bjork", artist),
    artist = ifelse(artist == blue_oyster_cult, "Blue Oyster Cult", artist),
    artist = ifelse(artist == husker_do, "Husker Do", artist),
    artist = ifelse(artist == bjork_brodsky, "Bjork & Brodsky Quartet", artist),
    artist = ifelse(artist == slagsmalsklubben, "Slagsmalsklubben", artist),
    artist = ifelse(artist == bjork_sugarcubes, "Bjork (Sugarcubes)", artist)
  )
```


Using the Reed College jukebox data, what are the top 10 artists played during
the "graveyard shift" during the academic year? Define

* the "graveyard shift" as midnight to 8am
* the academic year as September through May (inclusive)

```{r, echo=TRUE, message=TRUE, warning=TRUE, fig.width=8, fig.height=4.5}
graveyard <- jukebox %>%
  mutate(date_time = parse_date_time(date_time, "a b d HMS Y")) %>%
  separate(date_time, c("year", "month", "day", "hour", "minute", "seconds")) %>%
  select(year, month, day, hour, minute, artist, track)

schoolyard <- graveyard %>%
  transform(month=as.numeric(month)) %>%
  transform(hour=as.numeric(hour)) %>%
  filter(!month %in% c(6, 7, 8)) %>%
  filter(!hour<8) %>%
  group_by(artist) %>%
  tally() %>% 
  rename(playcount = n) %>%
  filter(!playcount < 894) 

ggplot(schoolyard, aes(x=fct_reorder(artist, playcount), y=playcount)) +
  geom_bar(stat = "identity", fill = "violetred4") +
  coord_flip()+
  labs(title="Most Popular Artists during the Graveyard Shift", x='Artist', y='Play Count') +
  theme_minimal()


```





